---
title: "rwr-cv-example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rwr-cv-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## 1. Introduction
This document descibes the usage of RWR_CV from the creation of a multiplex network from scratch to the parsing of RWR_CV's output. 

## 2. Setup
In order to run `RWR_CV`, we will need to generate an `mpo` object from the `RWRtoolkit::RWR_make_multiplex`.  An flist is first required to create an `mpo` object. We can generate random graphs using `igraph::erdos.renyi.game`. 

```{r setup}
library(RWRtoolkit)
library(igraph)
```

# https://stackoverflow.com/questions/34535155/figure-size-in-r-markdown
```{r Fig1, echo=FALSE, fig.height=5, figwidth=15}
set.seed(42)
layer1 <- igraph::erdos.renyi.game(5, 1/4)
layer2 <- igraph::erdos.renyi.game(20, 1/8)
layer3 <- igraph::erdos.renyi.game(10, 1/3)

par(mfrow=c(1,3), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
igraph::plot.igraph(layer1)
igraph::plot.igraph(layer2)
igraph::plot.igraph(layer3)
```

Once the networks have been generated, we can save these as an edge list and generate an flist for `RWR_make_multiplex`. 

```{r}


```