---
title: "Using RWR-LOE To Explore Knockout Pairs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using RWR-LOE To Explore Knockout Pairs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Background
`CITE` et al. have demonstrated phenotypic differences though two double knockout experiments: {FAE1, FAD2} and {FAE1, ROD1}. The {FAE1, FAD2} knockout experiment obtained a 90% accumulation of oleic acid, however, overall seed yield was reduced and growth stunted. Conversely, the {FAE1, ROD1} experiment saw a 60% accumulation of oleic acid but did not see stunted growth or reduction of seed yeild.  

Though both knockout experiments result in a greater accumulation of oleic acid, the authors noted that there was not immediately apparent explanation as to what could have caused the drastic difference in phenotypes.  `<<CITE THIS>>`

# Setup
```{r setup}
library(RWRtoolkit)
library(gprofiler2)
```

## Load Networks
Load Previously defined R libraries from the RWR toolkit Kbase networks: 
### Expression/Regulation Multiplex


| **Const** | Expression_Regulation_AT_d0.5_v01.RData |
| :---------------- | :------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| **Version** | 0.1 |
| **Description** | Contains the following network layers with unweighted edges and delta value = 0.5:<br />Coexpression Gene-Atlas (GA) version 0.3 (AT-UU-GA-01-AA-01)<br />PEN-Diversity (PX) version 0.1 (AT-UU-PX-01-AA-01)<br />Regulation-ATRM (RE) version 0.3 (AT-UU-RE-00-AA-01) |

Network variables will be: 

- `nw.mpo`:     The multiplex object
- `nw.adj`:     The supra-adjacency matrix
- `nw.adjnorm`  The normalized supra-adjacency matrix

NOTE: The below assumes that you've cloned the [RWRtoolkit-data](https://github.com/dkainer/RWRtoolkit-data) into an adjacent directory

```{r}
regulationNetworkPath <- '~/Documents/tools/RWRtoolkit-data/Expression_Regulation_AT_d0.5_v01.RData'
load(regulationNetworkPath)
```

Next, we can generate our gene lists to act as seeds for our separate LOE runs, with our genes converted to their corresponding TAIR IDs:  


| **Gene Name** | **TAIR ID**  |
|:----------------|:----------------|
| FAE1  | AT4G34520 | 
| FAD2  | AT3G12120 |
| ROD1  | AT3G15820 |

```{r}
write_genelist_to_file <- function(genelist, setlist, filename){
  df <- data.frame(setlist, genelist)
  write.table(df, filename, sep='\t', row.names = FALSE, quote= FALSE)
}
## FAE1 FAD2
fae1fad2List <-c('AT4G34520','AT3G12120') 
setlist1 <- c('set1','set1')
fea1fad2GenesetPath <- './genelist_fae1_fad2.tsv'
write_genelist_to_file(fae1fad2List, setlist1, fea1fad2GenesetPath)

## FAE1 ROD1
fae1rod1List <- c('AT4G34520', 'AT3G15820')
setlist2 <- c('set2','set2')
fae1rod1GenesetPath <- './genelist_fae1_rod1.tsv'
write_genelist_to_file(fae1rod1List, setlist2, fae1rod1GenesetPath)
system("pwd")
```

# Differential Ranking
To implement differential ranking, we must run RWR_LOE twice: once with the `fae1fad2GenesetPath` and once with the `fae1rod1GenesetPath`. By doing so, we will obtain the genes most highly ranked for each. 


# Exploring Topological Differences

We can extract rankings of the nearest gene nodes around our seed genes by running RWR LOE on our network and supplied seed genes. By supplying an integer to `cyto` we can then produce a cytoscape session with the ranked genes. 

Note the following commands assume a cytoscape is open. 
## Cytoscape

```{r echo=FALSE}
fea1fad2_loeOutput <- RWRtoolkit::RWR_LOE(data = regulationNetworkPath, 
                                            seed_geneset = fea1fad2GenesetPath, 
                                            outdir = './fae1fad2', 
                                            ) 

fae1rod1_loeOutput <- RWRtoolkit::RWR_LOE(data = regulationNetworkPath, 
                                            seed_geneset = fae1rod1GenesetPath, 
                                            outdir = './fae1rod1', 
                                            ) 
```

```{r eval=FALSE}
cyto_num <- 200
fea1fad2_loeOutput <- RWRtoolkit::RWR_LOE(data = regulationNetworkPath, 
                                            seed_geneset = fea1fad2GenesetPath, 
                                            outdir = './fae1fad2', 
                                            cyto = cyto_num) 

fae1rod1_loeOutput <- RWRtoolkit::RWR_LOE(data = regulationNetworkPath, 
                                            seed_geneset = fae1rod1GenesetPath, 
                                            outdir = './fae1rod1', 
                                            cyto = cyto_num) 
```


With the newly created cytoscape networks, three separate sub-networks can then be generated using the difference and intersect methods, from which ClueGO can be run on three distinc subnetworks. 


## R Code

Conversely, we can do the same experiment via code by taking the differences and intersection of the output ranked gene sets. 

### Setup
```{r}
# Read output files from corresponding directories
fae1fad2RankDF <- read.table('./fae1fad2/RWR-LOE_set1_default.ranks.tsv', header = TRUE)
fae1rod1RankDF <- read.table('./fae1rod1/RWR-LOE_set2_default.ranks.tsv', header = TRUE)

# Extract top 200 Genes: 
fae1fad2RanksTop200 = fae1fad2RankDF[fae1fad2RankDF['rank'] <= 200, ]
fae1rod1RanksTop200 = fae1rod1RankDF[fae1rod1RankDF['rank'] <= 200, ]

# Get Nodes and Append Seeds
fae1fad2GenesTop200 = fae1fad2RanksTop200['NodeNames']
fae1rod1GenesTop200 = fae1rod1RanksTop200['NodeNames']

# Append the seed nodes as they are not ranked within the output sets. 
fae1fad2Data <- append(fae1fad2GenesTop200@.Data[[1]], c('AT4G34520', 'AT3G12120'))
fae1rod1Data <- append(fae1rod1GenesTop200@.Data[[1]], c('AT4G34520', 'AT3G15820'))

# Get Set Differences and Intersection: 
fad2Diff <- setdiff(fae1fad2Data, fae1rod1Data)
rod1Diff <- setdiff(fae1rod1Data, fae1fad2Data)
fae1Intersection <- intersect(fae1fad2Data, fae1rod1Data)
```

### Go Enrichment
We now have three differing sets. One for genes only from the fae1fad2 network, one for genes only from the fae1rod1 network, and one for genes shared by both networks. Given our three entirely unique sets, we can now query a database to extract GO Enrichments using `gprofiler2`: 

```{r}
gostres_fad2 <- gost(query=fad2Diff, organism='athaliana')
gostres_rod1 <- gost(query=rod1Diff, organism='athaliana')
gostres_intersection <- gost(query=fae1Intersection, organism='athaliana')
```

### Results Analysis
#### FAD2 Difference Genes

```{r}
gostres_fad2$result[gostres_fad2$result['significant']==TRUE, ][ , c('p_value',  'term_name')]
```
#### ROD1 Difference Genes
ROD1 did not have any meaningful results. 
```{r}
gostres_rod1$result
```
#### Intersection
```{r}
gostres_intersection$result[gostres_intersection$result['significant']==TRUE, ][, c('p_value', 'term_name')]
```