---
title: "rwr-cv-example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rwr-cv-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## 1. Introduction
This document descibes the usage of RWR_CV from the creation of a multiplex network from scratch to the parsing of RWR_CV's output. 

## 2. Setup
In order to run `RWR_CV`, we will need to generate an `mpo` object from the `RWRtoolkit::RWR_make_multiplex`.  An flist is first required to create an `mpo` object. We can generate random graphs using `igraph::erdos.renyi.game`. 

```{r setup}
library(RWRtoolkit)
library(igraph)
```

### a. Network Layer Generation
```{r Fig1, echo=FALSE, fig.height=15, figwidth=5}
set.seed(42)
layer1 <- igraph::erdos.renyi.game(5, 1/4)
layer2 <- igraph::erdos.renyi.game(20, 1/8)
layer3 <- igraph::erdos.renyi.game(10, 1/3)

igraph::plot.igraph(layer1)
igraph::plot.igraph(layer2)
igraph::plot.igraph(layer3)
```

Once the networks have been generated, we can save these as an edge list
```{r}

## Function to convert edgelists from graphs and save to file. 
writeEdgelistToFile <- function(filename, layer){
    # Create Edgelist and rename columns 
    edgelist <- data.frame(as_edgelist(layer))
    colnames(edgelist) <- c('node1', 'node2')

    # Generate Weight list of all 1s (i.e. to make network unweighted)
    weights <- rep(1, nrow(edgelist)) 
    edgelist$weights <- weights    

    # save to file
    write.table(edgelist, filename, sep='\t', row.names = FALSE, quote= FALSE)
}

## Writing Layers to File
writeEdgelistToFile('layer1.tsv', layer1)
writeEdgelistToFile('layer2.tsv', layer2)
writeEdgelistToFile('layer3.tsv', layer3)
```
### b. `flist` and Multiplex Generation

To create the multiplex, first we need an flist (i.e. a file containing the file paths, short names, and grouplist ) pointing to our saved files and create our multiplex. We use 1s for all group values because we are making a node-to-node multiplex (i.e. treating all nodes as the same type such as gene to gene). 

```{r}
# Creating the FLIST
fileList <- c('layer1.tsv', 'layer2.tsv', 'layer3.tsv')
shortNameList <- c('L1', 'L2', 'L3')

groupList <- c(1, 1, 1)   
flistDF <- data.frame(fileList, shortNameList, groupList)
write.table(flistDF,'flistDF.tsv', sep='\t', col.names = FALSE, row.names = FALSE, quote = FALSE)

multiplexNetworkPath <- './toy_network.Rdata'

RWRtoolkit::RWR_make_multiplex(
   flist = './flistDF.tsv',
   output = './toy_network.Rdata',
   verbose = TRUE
 )
```


### c. Creating Goldset Genes

We can arbitrarily create a goldset of "functionally related" genes. We can designate genes 1, 2, 3, 4 to be in our original gold set given that we know these genes are highly connected within the previously created arbitrary layers (Typically, gold set genes are those annotated by a similar GO term or exist within the same biochemical pathway). 

```
1 goldSet
2 goldSet
3 goldSet
4 goldSet
```

We can then use our genes 1, 2, 3, 4 as seeds within our network to see how 

```{r}
goldsetGenes <- c(1, 3, 2, 4)
goldsetSetIDs <- c('goldSet', 'goldSet', 'goldSet', 'goldSet')
goldsetDF <- data.frame(goldsetSetIDs, goldsetGenes)

goldsetPath <- './goldsetCV.tsv'
write.table(goldsetDF, goldsetPath, row.names = FALSE, col.names = FALSE, sep='\t', quote=FALSE)
```

## Running RWR_CV

With our multiplex and gold set genes in place, we can run `RWR_CV`. First, let's run `loo` on the gold set genes to see how well they perform: 

```{r}
goldsetCV_looOutput <- RWRtoolkit::RWR_CV(
    dataPath = multiplexNetworkPath,
    genesetPath = genesetPath,
    method = 'loo', 
    restart = 0.7, 
    tau=c(1,1,1),
    verbose=TRUE
)

goldsetCV_looOutput$medianranks
```

Above, we can see the `$medianranks` output. For each node, the median value is taken across all full ranks (generated from running loo on each of the 4 seed genes).  All data can be seen by accessing the `$fullranks` of the output data. Metrics for each individual node can be found by accessing the `$metrics` data, and metrics for the entire network as a whole can be found under the `$summary` tag. 

