---
title: "Using RWR CV See Seed Set Connectivity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using RWR CV See Seed Set Connectivity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


Given some set of seeds: 

GABAergicSeeds  
serotonergicSeeds
From Sullivan et al. `<<CITE KYLE @ GRIN PAPER>>`'s work with the GRIN, we wish to test how closely functionally ranked some gene sets are: 


```{r setup}
library(RWRtoolkit)

humanMultiplexURL <- 'https://github.com/sullivanka/GRIN/blob/main/test/suicide_weighted_Multiplex_0.5Delta.RData?raw=true'
humanMultiplexPath <- url(grinMultiplexURL)

load(humanMultiplexPath)
```


Generating the gene sets from scratch: 
```{r}
GABAergicSeedGenes <- c( 'ENSG00000171132', 'ENSG00000109189', 'ENSG00000120907', 'ENSG00000169750', 'ENSG00000171862', 'ENSG00000188822',
'ENSG00000011677','ENSG00000022355','ENSG00000163812','ENSG00000188603','ENSG00000102287','ENSG00000133703','ENSG00000109158','ENSG00000169760',
'ENSG00000118432','ENSG00000135312','ENSG00000168748','ENSG00000164129','ENSG00000169992','ENSG00000115896','ENSG00000163285','ENSG00000146242',
'ENSG00000007516','ENSG00000113327','ENSG00000174576','ENSG00000106462','ENSG00000136854','ENSG00000109572','ENSG00000182256','ENSG00000170759',
'ENSG00000186297','ENSG00000149295','ENSG00000173805','ENSG00000145864','ENSG00000151834','ENSG00000214285','ENSG00000160446','ENSG00000010322',
'ENSG00000145863','ENSG00000152910','ENSG00000128271','ENSG00000139182','ENSG00000154822','ENSG00000122733','ENSG00000006128','ENSG00000185666',
'ENSG00000180914','ENSG00000163485','ENSG00000157103' )
seedGeneSets <- GAGAergicrep('GAGAergic_GO', length(GABAergicSeedGenes))

seedGeneDF <- data.frame('set'=rep('GAGAergic_GO', length(GABAergicSeedGenes)), 'seed'=GABAergicSeedGenes)

filenameGoldset <- './goldset_gaba.tsv')
write.table(seedGeneDF, filenameGoldset, sep='\t', row.names = FALSE, quote= FALSE)
```

## Sanity Check: 
Let's instead run RWR CV with a set of seed genes of the same size as above, but with 2/3 of our genes being randomly sampled from our network and replace the first N genes from our above `seedGeneDF` with these randomly sampled ones: 

```{r}
set.seed(42)
newlySampleGenesPct <- 2/3
newlySampleGeneCount <- floor(length(GABAergicSeedGenes) * newlySampleGenesPct)

randomSample <- sample(nw.mpo$Pool_of_Nodes, replace=FALSE, size=newlySampleGeneCount)

## Replace: 
seedGeneDF$seed[1:length(randomSample)] <- randomSample

filenameRandomSample <- './partiallyRandomSampleSeeds.tsv'
write.table(seedGeneDF, filenameRandomSample, row.names=FALSE, quote=FALSE)
```

# Run RWR CV 
```{r}
goldsetCV_looOutput <- RWRtoolkit::RWR_CV(
    dataPath = humanMultiplexPath,
    genesetPath = filenameGoldset,
    method = 'loo', 
    restart = 0.7, 
    tau=c(1,1,1),
    verbose=TRUE
)

goldsetCV_looOutput$medianranks
```



```{r}
goldsetCV_looOutput <- RWRtoolkit::RWR_CV(
    dataPath = humanMultiplexPath,
    genesetPath = filenameRandomSample,
    method = 'loo', 
    restart = 0.7, 
    tau=c(1,1,1),
    verbose=TRUE
)

goldsetCV_looOutput$medianranks
```